---
title: "Final Project - Is there an association with age and higher levels of HbA1c?"
Author: Andy Higuera
output:
  html_document:
    df_print: paged
---

# Introduction/Setup

Data dictionary: 
ID = unique identifier to patient (integer)
No_Pation = Patient ID (integer)
Gender = Gender of patient, either (string):
    Male (M)
    Female (F)
Age = Age of patient (integer)
Urea = Urea levels of patient (double)
Cr = Creatine levels of patient (double)
HbA1c = HbA1c levels of patient (double)
Chol = Cholesterol levels of patient (double)
TG = Triglyceride levels of patient (double)
HDL = HDL levels of patient (double)
LDL = LDL levels of patient (double)
VLDL = VLDL levels of patient (double)
BMI = BMI of patient (double)
CLASS = Type of diabetic class for patient, either (string):
  Non-diabetic (N)
  Predict-Diabetic (P)
  Diabetic (Y)

```{r include=FALSE}
library(tidyverse)
diabetes_data = read_csv("data/Dataset of Diabetes.csv")
# The variable below excludes duplicate patient ID, but keeps the first occurrence. 
diabetes_data2 = diabetes_data[!duplicated(diabetes_data$No_Pation), ]

```

# Figure 1.
```{r}
plot1 = 
  ggplot(diabetes_data2, aes(x=AGE, y=HbA1c, col=CLASS)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Relationship between age and HbA1c levels in patients.")
plot1
```


# Figure 2. In order to check for the distribution of the graph. We must look at it through a histogram. 
```{r}
plot13 = 
  ggplot(diabetes_data2, aes(x=AGE)) +
  geom_histogram()
plot13
```


# Figure 3. Statistical testing: Since the graph is normally distributed we must use the Pearson's method. 

```{r}
stat_test1 = cor(diabetes_data2$AGE, diabetes_data2$HbA1c, method="pearson")
stat_test1
```
## Give the result of the pearson statistical test we can 


# Figure 4. Confounder 1: BMI levels of patient. 
```{r}
confounder1 = diabetes_data2 %>%
  select(AGE, HbA1c, BMI, CLASS)

confounder1$BMI = 
  ifelse(confounder1$BMI<18.5, "Underweight",
         ifelse(confounder1$BMI<25, "Normal weight", "Obese"))

confounder1 %>%
  ggplot(aes(x = AGE, y = HbA1c, col=BMI)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ BMI) +
  labs(title = "Patient age and HbA1v compared with BMI levels")
```


# Confounder 2: Seperated LDL by ranges Optimal through Very high. Ranges referenced from HealthTab - Reference Ranges. (n.d.). https://www.healthtab.com/en/public_site/reference_ranges.
```{r}
confounder2 = diabetes_data2 %>%
  select(AGE, HbA1c, LDL, CLASS)

confounder2$LDL_label <- ifelse(confounder2$LDL < 2.6, "Optimal", 
                                ifelse(confounder2$LDL < 3.4, "Near optimal",
                                       ifelse(confounder2$LDL < 4.2, "Borderline high",
                                              ifelse(confounder2$LDL < 4.9, "High", "Very High"))))
```


# Figure 5. Confounder 2 LDL levels of patients
```{r}
confounder2 %>% 
  ggplot(aes(x = AGE, y = HbA1c, col=LDL_label)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ LDL_label) +
  labs(title = "Patients age and HbA1c levels compared with LDL levels")
```

